\documentclass{article}
\usepackage{amsmath, amssymb, mathtools, verbatim}
\usepackage[margin=0.5in]{geometry}
\setlength{\parindent}{0pt}
\renewcommand{\t}[1]{\text{#1}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\sumx}[2]{\sum\limits_{#1}^{#2}}
\newcommand{\ord}{\text{ord}}
\newcommand{\frp}{\text{FRP}}
\newcommand{\R}{\Rightarrow\,}
\newcommand{\B}[1]{\textbf{#1}}
\newcommand{\I}{\bigcap}
\newcommand{\U}{\bigcup}
\newcommand{\V}{\vert}
\newcommand{\F}[1]{\lfloor#1\rfloor}
\newcommand{\C}[1]{\lceil#1\rceil}
\renewcommand{\t}[1]{\text{#1}}
\newcommand{\lra}[1]{\langle#1\rangle}
\newcommand{\len}[1]{\vert#1\vert}
\newcommand{\num}{\text{num}}
\newcommand{\game}[3]{\begin{array}{@{}*6r}(#1, & #2, & #3)\end{array}}
\newcommand{\gcong}{\cong\:}
\begin{document}
\begin{center}
  21--499 \\
  Kartik Sabharwal (ksabharw)
\end{center}
\paragraph{Describing the Game}\mbox{}\\
This is a game for two players, who take turns making moves. \\
We have two points $P_1$ and $P_2$ in a plane, connected by an edge. \\
Conisder a function $s : \{P_1, P_2\} \rightarrow \N$. \\
Initially, $s(P_1) \geq s(P_2)$. \\
There are $s(P_1)$ stones at $P_1$ and $s(P_2)$ stones at $P_2$. \\
A player can move stones from $P_1$ to $P_2$, but not from $P_2$ to $P_1$,
but the invariant $s(P_1) \geq s(P_2)$ should always be maintained. \\
If a player starts her turn with $s(P_1) = s(P_2)$, then she loses. \\
What is the nimber of this game? \\
It ought to be $\frac{s(P_1) - s(P_2)}{2}$.

\paragraph{A Modified Version}\mbox{}\\
Suppose that a few things change from before:
\begin{enumerate}
  \item Either $s(P_1) > s(P_2)$ or $s(P_2) > s(P_1)$ initially
  \item We can move stones from $P_1$ to $P_2$ or $P_2$ to $P_1$
  \item Suppose at the beginning of some turn, $\len{s(P_1)-s(P_2)} = d$, and
    at the end of that turn $\len{s(P_1) - s(P_2)} = d'$. \\
    Then it must be that $d' < d$.
  \item If a player starts with $s(P_1) = s(P_2)$ then the player loses.
\end{enumerate}
How does that change the nimber of the game? \\
If we start with $s(P_1) = 10$ and $s(P_2) = 5$ then $G \cong *2$. \\
The nimber ought to be $\len{\frac{s(P_1) - s(P_2)}{2}}$.

\paragraph{Three Points}\mbox{}\\
Our point set $\mathbb{P}$ is now $\{P_1, P_2, P_3\}$, with directed edges
$(P_1, P_2)$ and $(P_2, P_3)$. \\
Start with the configuration $s(P_1) > s(P_2) > s(P_3)$. \\
Maintain the following invariants:
\begin{enumerate}
  \item $s(P_1) \geq s(P_2)$
  \item $s(P_2) \geq s(P_3)$
\end{enumerate}
If a player cannot make any moves, then the player loses. \\
This means we can describe a game of this form with a tuple
$(a,b,c)$ which means $s(P_1) = a, s(P_2) = b, s(P_3) = c$ initially. \\
Examples:
\begin{enumerate}
  \item $(3, 2, 1) \cong *0$ 
  \item $(3, 2, 2) \cong *0$
  \item $(3, 3, 0) \cong *1$ 
  \item $(3, 3, 1) \cong *1$
  \item $(3, 3, 2) \cong *0$
  \item $(4, 1, 1) \cong *1$ 
  \item $(4, 2, 0) \cong *0$ 
  \item $(4, 2, 1) \cong *0$
  \item $(4, 2, 2) \cong *1$
  \item $(4, 3, 0) \cong *1$
  \item $(4, 3, 1) \cong *0$
  \item $(4, 4, 0) \cong *2$
  \item $(5, 1, 1) \cong *2$
  \item $(5, 2, 0) \cong *0$
  \item $(5, 2, 1) \cong *1$
  \item $(5, 3, 0) \cong *0$
  \item $(6, 1, 1) \cong *2$
  \item $(6, 2, 0) \cong *1$
\end{enumerate}
\paragraph{A New Three Point Situation}\mbox{}\\
The point set $\mathbb{P}$ is again $\{P_1, P_2, P_3\}$ where
the directed edge set $\mathbb{E}$ is $\{(P_2, P_1), (P_2, P_3)\}$. \\
The initial state is of the form $s(P_1) = 0, s(P_2) = n, s(P_3) = 0$. \\
The rest of the rules are as before. \\
What is the nimber for some given value of $n$? \\
According to the program:
\begin{itemize}
  \item $(0, 1, 0) \cong *0$
  \item $(0, 2, 0) \cong *1$
  \item $(0, 3, 0) \cong *0$
  \item $(0, 4, 0) \cong *2$
  \item $(0, e, 0) \cong *1$ for $e = 2i$ where $i \in \N, i > 2$
  \item $(0, o, 0) \cong *0$ for $o = 2i - 1$ where $i \in \N, i > 2$
\end{itemize}

\paragraph{A helpful identity to prove}\mbox{}\\
\begin{equation*}
  \game{l}{m}{r} \gcong \game{l + i}{m + i}{r + i}
\end{equation*}

\newpage
\paragraph{Three Point Situation, 1-Position Proof, Case 8}\mbox{}\\
We want to prove the following proposition:
\begin{equation*}
  P(n) : \game{0}{6n + 8}{0} \gcong *1 \enspace \text{where} \enspace n \geq 0
\end{equation*}
We will use strong induction to prove the above statement.

\bigskip
\textbf{Base Cases:} \\
Using our program, we have verified that $P(n)$ holds for
$n \in \{0, 1, 2, 3\}$. \\
The fact that $\game{0}{2}{0} \gcong *1$ will also be useful.

\bigskip
\textbf{Induction Hypothesis:} \\
Assume for some $k \geq 1$ that 
$\forall \: n < k, \: P(n) \: \text{holds}$.

\bigskip
\textbf{Induction Step:} \\
We need to prove that $P(k)$ holds. In other words, we need to show:
\begin{equation*}
  \game{0}{6k + 8}{0} \gcong *1
\end{equation*}

\bigskip
\textit{Simplifying the problem,}\\
To show that $\game{0}{6k+8}{0}$ has nimber $1$, it suffices to show
the following:
\begin{enumerate}
  \item In one move, we can reach a state with nimber $0$
  \item In one move, we cannot reach a state with nimber $1$
\end{enumerate}

\bigskip
\textit{Looking at the initial state's children,}\\
Notice that in one move from the original state we can reach any state
of the form:
\begin{equation*}
  \game{m}{6k+8-m}{0} \text{where} \: 1 \leq m \leq 3k + 4 
  \end{equation*}
the remaining states that we can reach from the original state are of
the form:
\begin{equation*}
  \game{0}{6k+8-m}{m} \text{where} \: 1 \leq m \leq 3k + 4
\end{equation*}
Since the latter form is symmetric to the former, focusing our analysis
on the former is enough to complete the proof.

\bigskip
\textit{When} $m = 3k + 4$:\\
Observe that when $m = 3k + 4$, we obtain the state
$\game{3k+4}{3k+4}{0}$ which clearly has nimber $0$,
so we have proved the first fact we need. \\

\bigskip
\textit{When} $m$ \textit{is odd and} $1 \leq m \leq 2k + 2$:\\
When $m$ is an odd number satisfying $1 \leq m \leq 2k + 2$,
notice that
\begin{equation*}
  \game{m+1}{3k+4-\frac{m+1}{2}}{3k+4-\frac{m+1}{2}} \gcong *0
\end{equation*}
Work one step backwards from this to get
\begin{equation*}
  \game{m}{3k+5-\frac{m+1}{2}}{3k+4-\frac{m+1}{2}} \gcong *1
\end{equation*}
Observe that
\begin{equation*}
  \game{m}{6k+8-m}{0} \xrightarrow{\text{one move}}
  \game{m}{3k+5-\frac{m+1}{2}}{3k+4-\frac{m+1}{2}}
\end{equation*}
It follows that the minimum excluded value of the nimbers of
all the child states of $\game{m}{6k+8-m}{0}$ cannot be $1$, and
thus the nimber of this state cannot be 1.

\newpage
\textit{When} $m$ \textit{is even and} $1 \leq m \leq 2k + 2$:\\
When $m$ is an even number satisfying $1 \leq m \leq 2k + 2$,
observe that 
\begin{equation*}
  \game{m}{6k+8-m}{0} \xrightarrow{\text{one move}}
  \game{m}{6k+8-2m}{m}
\end{equation*}
We can replace $m$ with $2i$ where $1 \leq i \leq k + 1$ to get
\begin{equation*}
  \game{2i}{6k+8-2i}{0} \xrightarrow{\text{one move}}
  \game{2i}{6k+8-4i}{2i}
\end{equation*}
In the resultant state, the difference between the left and middle
values, also the difference between the right and middle values, is
\begin{equation*}
  6(k - i) + 8
\end{equation*}
Which intuitively suggests that
\begin{equation*}
  \game{2i}{6k+8-4i}{2i} \gcong \game{0}{6(k - i) + 8}{0} \:
  \text{for $1 \leq i \leq k + 1$}
\end{equation*}
and although we lack a rigorous argument for this we will assume it is
true. \\
Our knowledge that $\game{0}{2}{0} \gcong *1$ and also our induction
hypothesis, tell us:
\begin{equation*}
  \game{0}{6(k - i) + 8}{0} \gcong *1 \:
  \text{for $1 \leq i \leq k + 1$}
\end{equation*}
Consequently, $\game{2i}{6k+8-4i}{2i} \gcong *1$. \\
It follows that the minimum excluded value of the nimbers of
all the child states of $\game{m}{6k+8-m}{0}$ cannot be $1$, and
thus the nimber of this state cannot be 1.

\bigskip
\textit{When} $m \geq 2k + 3$:\\
If $m \geq 2k + 3$, it is clear that
\begin{equation*}
  \game{m}{m}{6k+8-2m} \gcong *0
\end{equation*}
Working backwards from this, we get that
\begin{equation*}
  \game{m}{m+1}{6k+7-2m} \gcong *1
\end{equation*}
We know that
\begin{equation*}
  \game{m}{6k+8-m}{0} \xrightarrow{\text{one move}}
  \game{m}{m+1}{6k+7-2m}
\end{equation*}
It follows that the minimum excluded value of the nimbers of
all the child states of $\game{m}{6k+8-m}{0}$ cannot be $1$, and
thus the nimber of this state cannot be 1.

\bigskip
\textit{Conclusion}:\\
$1$ must be the mex of the nimbers of the children of $\game{0}{6k+8}{0}$,
so we have proved that $\game{0}{6k+8}{0} \gcong *1$.
\newpage

\paragraph{Three Point Situation, 1-Position Proof, Case 6} \\
Assume the same inductive `skeleton' and reasoning style as the previous case.
This proof is mainly an illustration of the differences between the two cases:

\bigskip
Consider the initial position
\begin{equation*}
  \game{0}{6n+6}{0} \: \text{where} \: n \geq 0
\end{equation*}

\medskip
The first move will take the initial position to a position of the form
\begin{equation*}
  \game{m}{6n+6-m}{0} \: \text{where} \: 1 \leq m \leq 3n+3
\end{equation*}

\bigskip
\underline{If $m = 3n$:} \\
This is the state $\game{3n}{3n}{0}$ which clearly has nimber $0$.

\bigskip
\underline{If $m \leq 2n + 2$ and $m$ is odd:} \\
Then, consider the following move sequence,
\begin{align*}
  & \game{m}{6n+6-m}{0} \\
  \xrightarrow{\text{one possible child}} \quad & 
  \game{m}{3n+4-\frac{m+1}{2}}{3n+3-\frac{m+1}{2}} \\
  \xrightarrow{\text{only child}} \quad & 
  \game{m+1}{3n+3-\frac{m+1}{2}}{3n+3-\frac{m+1}{2}} \\
\end{align*}

\bigskip
\underline{If $m \leq 2n + 2$ and $m$ is even:} \\
Stuff.

\bigskip
\underline{If $m \geq 2n + 3$:} \\
Stuff.

\bigskip
Which provides us with everything we need to fill in the proof
skeleton.
\paragraph{Three Point Situation, 1-Position Proof, Case 10}
\end{document}
